---
title: "stampr analysis"
author: "Colin Robertson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Introduction to analysis of moving polygons in space-time using the `stampr` package.

```{r}
library(stampr)
library(sp)
library(igraph)
data('katrina')
class(katrina)
head(katrina@data)
```

## Simple two-time period change detection using overlay

```{r tidy=TRUE}
T1 <- katrina[1,]
T2 <- katrina[2,]
plot(T1, col=NA, border="green", lwd=2, xlim=c(min(c(bbox(T1)[1,1], bbox(T2)[1,1])), max(c(bbox(T1)[1,2], bbox(T2)[1,2]))), ylim=c(min(c(bbox(T1)[2,1], bbox(T2)[2,1])), max(c(bbox(T1)[2,2], bbox(T2)[2,2]))))
plot(T2, col=NA, border="red", add=TRUE, lwd=2)
```

The change we want to detect is the proportion of overlap, green only, and red only areas, representing stability, contraction and expansion events respectively.

```{r tidy=TRUE}
ch <- stamp(T1, T2, dc=0, direction=TRUE, distance=TRUE)
stamp.map(ch)
head(ch@data)
```

## Multiple time period polygon change analysis using overlay
```{r tidy=TRUE}
T3 <- katrina[3,]
ch2 <- stamp(T2, T3, dc=0, direction=TRUE, distance=TRUE)
head(ch2@data)
```

## Multiple time period polygon change analysis using overlay
```{r tidy=TRUE}
outEvents <- list()
lSum <- 1
for(i in 1:(nrow(katrina@data)-1)) {
  Ti <- katrina[i,]
  Ti_1 <- katrina[i+1,]
  ch <- stamp(Ti, Ti_1, dc=0, direction=TRUE, distance=TRUE)
  xx1 <- spChFIDs(ch, paste(i, "-", as.character((lSum):(lSum+length(ch)-1)), sep=""))
  lSum <- length(ch) + lSum
  outEvents[[i]] <- xx1 #@data
} 

outEvents <- do.call("rbind", outEvents)
```

## Get summaries from multiple change events
```{r tidy=TRUE}
grps <- row.names(outEvents) #get row names
outEvents$TGROUP <- substr(grps, 1, as.numeric(unlist(lapply(strsplit(grps, ''), function(x) which(x == '-'))))-1) #extract time period to distinguish change time periods
outEvents$STGROUP <- as.numeric(paste(outEvents$TGROUP,outEvents$GROUP, sep="")) #generate group ID for specific time periods

```

##Do some more graphing of the topological relationships
```{r tidy=TRUE}
outSTGroup <- stamp.stgroup.summary(outEvents)
df <- data.frame(from=outEvents$ID1,to=outEvents$ID2, stg = outEvents$STGROUP)
df <- df[complete.cases(df),]
df <- merge(outSTGroup, df, by.x="STGROUP", by.y="stg")
df$weight <- (df$aSTBL / df$AREA) * 10
df <- data.frame(from=df$from,to=df$to, weight = df$weight)
g <- graph_from_data_frame(df, directed = TRUE, vertices=df$ID1)
E(g)$weight <- df$weight 
plot(g, edge.width=E(g)$weight, layout=layout.fruchterman.reingold, edge.curved=TRUE, vertex.size=5, edge.arrow.mode='-')
```

Lets look at T=1:10, and T=11:20, and T=21:32 separately to see if we can see any changes in space-time structure...
```{r tidy=TRUE}
df2 <- df[1:10,]
g <- graph_from_data_frame(df2, directed = TRUE, vertices=df2$ID1)
E(g)$weight <- df2$weight
plot(g, edge.width=E(g)$weight, layout=layout.fruchterman.reingold, edge.curved=TRUE, vertex.size=5, edge.arrow.mode='-')
df2 <- df[11:20,]
g <- graph_from_data_frame(df2, directed = TRUE, vertices=df2$ID1)
E(g)$weight <- df2$weight
plot(g, edge.width=E(g)$weight, layout=layout.fruchterman.reingold, edge.curved=TRUE, vertex.size=5, edge.arrow.mode='-')
df2 <- df[21:32,]
g <- graph_from_data_frame(df2, directed = TRUE, vertices=df2$ID1)
E(g)$weight <- df2$weight
plot(g, edge.width=E(g)$weight, layout=layout.fruchterman.reingold, edge.curved=TRUE, vertex.size=5, edge.arrow.mode='-')

```

##Space-time Graph Clustering
